---
title: "Financial Risk Management"
output:
  html_notebook:
    toc: true
    toc_float: true
---

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r include=FALSE}
library(xts)
library(rmgarch)
library(fOptions)
library(openxlsx)
library(tidyverse)
library(lubridate)
library(SciViews)
library(ggplot2)
library(scales)
options(scipen = 999)
```

## Greek Letter Risks of Portfolio

### Load Portfolio and Yield Curve
```{r}
optionData <- read.xlsx("optiondata.xlsx", rows = c(5:25), detectDates = TRUE, sheet = "Option prices")
yieldCurve <- read.xlsx("optiondata.xlsx", sheet = "zero-coupon yield curve")

# Get SPX info
currentInfo <- read.xlsx("optiondata.xlsx", rows = c(1:3), cols = c(1:2), colNames = FALSE, sheet = "Option prices")

DATE <- as.Date(currentInfo[1, 2], origin = "1899-12-30") # Evaluation date 2020/10/30
SPOT <- currentInfo[2, 2] # Spot price of SPX on evaluation date
q <- currentInfo[3, 2] # Dividend yield of SPX

optionData
```

### Time to Expiry
```{r}
n <- nrow(optionData) # Number of option positions

# Calculate time to expiry in terms of days

# If the Expiration.Time column Open and we need to minus 6.5/24 
# If the Expiration.Time column Close and we do not need to minus 6.5/24 

time_to_expiry_days <- rep(0, n)

for (i in 1:n) {
  t <- optionData$Expiration.Time[i]
  d <- optionData$Expiration.Date[i]
  
  if (t == "Open") {
    time_to_expiry_days[i] <- d - DATE - (6.5/24)
  } else {
    time_to_expiry_days[i] <- d - DATE
  }
}

# Add time_to_expiry to optionData
optionData$time_to_expiry_days <- time_to_expiry_days
optionData$time_to_expiry_years <- time_to_expiry_days / 365 # Annualized time_to_expiry
optionData$time_to_expiry_days
```

### Interest Rates 
Linear Interpolation of interest rate
Rn = R1 + (R2 - R1) / (t2 - t1) * (tn - t1)
```{r}
rates <- rep(0, n)

# Rates for each position
for (i in 1:n) {
  tn <- optionData$time_to_expiry_days[i]
  
  # Search through yield curve to find appropriate parameters
  for (j in 1:nrow(yieldCurve)) {
    if (yieldCurve$days[j] >= tn) {
      t1 <- yieldCurve$days[j - 1]
      t2 <- yieldCurve$days[j]
      r1 <- yieldCurve$rate[j - 1]
      r2 <- yieldCurve$rate[j]
      break
    }
  }
  
  # Linear interpolation
  rates[i] <- r1 + (r2 - r1) * (tn - t1) / (t2 - t1)
}

# Add rates to optionData
optionData$Rate <- rates / 100 # rates in %
optionData$Rate
```

### Implied Volatility
```{r}
vol <- rep(0, n)

for (i in 1:n) {
  vol[i] <- GBSVolatility(price = optionData[i, "Price"], TypeFlag = "c", S = SPOT, X = optionData[i, "Strike.Price"], Time = optionData[i, "time_to_expiry_years"], r = optionData[i, "Rate"], b = optionData[i, "Rate"] - q)
}

optionData$Implied.volatility <- vol
optionData$Implied.volatility
```

### Compute d1 and d2
equation
R vectorization soooo gooood
```{r}
X <- optionData$Strike.Price
r <- optionData$Rate
sig <- optionData$Implied.volatility
t <- optionData$time_to_expiry_years

optionData$d1 <- ((ln(SPOT/X) + (r - q + 0.5 * sig ^ 2) * t)) / (sig * sqrt(t))
optionData$d2 <- optionData$d1 - sig * sqrt(t)
```

### Greeks
equations
```{r}
X <- optionData$Strike.Price
multiplier <- optionData$Multiplier
t <- optionData$time_to_expiry_years
r <- optionData$Rate
sig <- optionData$Implied.volatility
d1 <- optionData$d1
d2 <- optionData$d2

optionData$Delta <- multiplier * exp(-q * t) * pnorm(d1)
optionData$Gamma <- multiplier * exp(-q * t) * dnorm(d1) / (SPOT * sig * sqrt(t))
optionData$Theta <- multiplier * {-0.5 * SPOT * sig * exp(-q * t) * dnorm(d1) / sqrt(t) - r * X * exp(-r * t) * pnorm(d2) + q * SPOT * exp(-q * t) * pnorm(d1)}
optionData$Vega  <- multiplier * SPOT * exp(-q * t) * sqrt(t) * dnorm(d1)
optionData
```

### Portfolio Greeks
```{r}
portfolioGreeks <- data.frame(row.names = "Portfolio",
                              Delta = sum(optionData$Delta * optionData$Quantity),
                              Gamma = sum(optionData$Gamma * optionData$Quantity),
                              Theta = sum(optionData$Theta * optionData$Quantity),
                              Vega = sum(optionData$Vega * optionData$Quantity))
portfolioGreeks
```

### Portfolio value as SPX value changes
```{r}
# Calculate portfolio value for SPX values ranging from -20% to +20% of spot price
value <- c()
for (s in as.integer(0.8 * SPOT):as.integer(1.2 * SPOT)) {
  
  # Calculate value for each option position
  value_part <- rep(0, n)
  for (i in 1:n) {
    value_part[i] <- optionData[i, "Quantity"] * optionData[i, "Multiplier"] * 
      GBSOption(TypeFlag = "c", S = s, X = optionData[i, "Strike.Price"], Time = optionData[i, "time_to_expiry_years"], r = optionData[i, "Rate"], b = optionData[i, "Rate"] - q, sigma = optionData[i, "Implied.volatility"])@price
  }
  
  # Append value
  value <- c(value, sum(value_part))
}

```

### Data Visualization
```{r}
priceChange <- data.frame(SPX = seq(-20, 20, 40 / (as.integer(1.2 * SPOT) - as.integer(0.8 * SPOT))), 
                          Value = value)

priceChange %>% 
  ggplot(aes(x=SPX, y=Value/1000000)) +
  geom_line() +
  ggtitle("Portfolio value as a function of % change in stock index") +
  xlab("% change in stock index") +
  ylab("Portfolio value (million $)") +
  scale_y_continuous(labels = comma)
```

## Modeling SPX and VIX Index

### Load data
```{r}
SPX_history <- read.csv("SPX_History.csv")
VIX_history <- read.csv("VIX_History.csv")

SPX <- subset(SPX_history, select = c(Date,Close))
VIX <- subset(VIX_history, select = c(Date,Close))
SPX$Date <- as.Date(SPX$Date, format="%m/%d/%Y")
VIX$Date <- as.Date(VIX$Date, format="%m/%d/%Y")

# Merge dataframes
colnames(SPX)[colnames(SPX) == "Close"] <- "SPX"
colnames(VIX)[colnames(VIX) == "Close"] <- "VIX"
indexData <- merge(SPX, VIX, by = "Date")
```

### Compute Index Log Return
```{r}
n_idx <- nrow(indexData)
indexData$SPX.Return[2:n_idx] <- log(indexData$SPX[2:n_idx] / indexData$SPX[1:(n_idx - 1)])
indexData$VIX.Return[2:n_idx] <- log(indexData$VIX[2:n_idx] / indexData$VIX[1:(n_idx - 1)])

# Get the most recent 1000 observations
idx_DATE <- which(indexData$Date == DATE)
returns <- indexData[(idx_DATE - 999):idx_DATE, c("Date", "SPX.Return", "VIX.Return")]
```

### Model Log Returns using GARCH
```{r}
# Univariate GARCH specification prior to fitting
# Estimation of the parameters
spec1 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                    mean.model = list(armaOrder = c(0,0), include.mean = FALSE),
                    distribution.model = "norm")

spec2 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                    mean.model = list(armaOrder = c(1,0), include.mean = TRUE),
                    distribution.model = "norm")

# Fitting GARCH model
fitSPX <- ugarchfit(spec = spec1, data = returns[, "SPX.Return"])
fitVIX <- ugarchfit(spec = spec2, data = returns[, "VIX.Return"])

# GARCH coefficients
coef(fitSPX)
coef(fitVIX)
```

### Estimation of the DCC model
```{r}
multiSpec <- multispec(c(spec1, spec2))
dccSpec <- dccspec(multiSpec, dccOrder = c(1,1), model = "DCC", distribution = "mvnorm")
fitDcc <- dccfit(dccSpec,returns[, c("SPX.Return", "VIX.Return")])

# DCC coefficients
coef(fitDcc)
```

### Store Coefficients
```{r}
SPXcoef <- fitSPX@fit$coef 
SPXsigma <- fitSPX@fit$sigma  
SPXz <- fitSPX@fit$z          
VIXcoef <- fitVIX@fit$coef   
VIXsigma <- fitVIX@fit$sigma  
VIXz <- fitVIX@fit$z          
DCCcoef <- fitDcc@mfit$coef        
DCCQ <- fitDcc@mfit$Q
```

### Conditional Mean, Variance, and Correlation for the next date
equations
```{r}
# Conditional mean
SPXmean <- 0
VIXmean <- VIXcoef["mu"] + VIXcoef["ar1"] * returns[1000, "VIX.Return"]

# Conditional variance
SPXvariance <- SPXcoef["omega"] + SPXcoef["alpha1"] * returns[1000, "SPX.Return"]^2 + SPXcoef["beta1"] * SPXsigma[1000]^2
VIXvariance <- VIXcoef["omega"] + VIXcoef["alpha1"] * returns[1000, "VIX.Return"]^2 + VIXcoef["beta1"] * VIXsigma[1000]^2

# Conditional correlation
rhobar <- (1/1000) * SPXz %*% VIXz
q11 <- 1.0 + DCCcoef["[Joint]dcca1"] * (SPXz[1000]^2 - 1) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][1,1] - 1)
q22 <- 1.0 + DCCcoef["[Joint]dcca1"] * (VIXz[1000]^2 - 1) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][2,2] - 1)
q12 <- rhobar + DCCcoef["[Joint]dcca1"] * (SPXz[1000] * VIXz[1000] - rhobar) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][1,2] - rhobar)
rho <- q12 / sqrt(q11 * q22)
```

### Portfolio Expected Change, Variance, and Delta-Normal VaR (Value at Risk)
```{r}
delta <- portfolioGreeks$Delta
vega <- portfolioGreeks$Vega

# Portfolio expected change
portExpectedChange <- vega * (indexData[idx_DATE, "VIX"] / 100) * VIXmean

# Portfolio Variance
portVariance <- delta^2 * indexData[idx_DATE, "SPX"]^2 * SPXvariance + vega^2 * (indexData[idx_DATE, "VIX"] / 100)^2 * VIXvariance + delta * vega * indexData[idx_DATE, "SPX"] * (indexData[idx_DATE, "VIX"] / 100) * rho * sqrt(SPXvariance * VIXvariance)

# Portfolio standard deviation
portSD <- sqrt(portVariance)

# Portfolio delta-normal VaR
DNVaR <- -1.645 * (portExpectedChange - portSD)
DNVaR
```

## Filter Historical Simulation






