---
title: Financial Risk Management
author: BRIAN WU
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
    df_print: paged
---

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

## Introduction
This project ...
see source code at my [github repository](https://github.com/BriAnWuu/Financial_Risk_Management)

### Install Packages
All R packages below are on CRAN, and can be installed with `install.packages`.
```{r}
library(xts)
library(rmgarch)
library(fOptions)
library(openxlsx)
library(tidyverse)
library(lubridate)
library(SciViews)
library(ggplot2)
library(scales)
options(scipen = 999)
```

## Portfolio Overview

```{r include=FALSE}
optionData <- read.xlsx("data/optiondata.xlsx", rows = c(5:25), detectDates = TRUE, sheet = "Option prices")
yieldCurve <- read.xlsx("data/optiondata.xlsx", sheet = "zero-coupon yield curve")

# Get SPX info
currentInfo <- read.xlsx("data/optiondata.xlsx", rows = c(1:3), cols = c(1:2), colNames = FALSE, sheet = "Option prices")

DATE <- as.Date(currentInfo[1, 2], origin = "1899-12-30") # Evaluation date 2020/10/30
SPOT <- currentInfo[2, 2] # Spot price of SPX on evaluation date
q <- currentInfo[3, 2] # Dividend yield of SPX
```

### Portfolio
```{r}
optionData
```

### Yield Curve
```{r}
yieldCurve
```

### Important Information (Parameters that will be used throughout this Project)
```{r}
DATE # Evaluation date 2020/10/30
SPOT # Spot price of SPX on evaluation date
q # Dividend yield of SPX
```


## Greek Letter Risks of Portfolio

### Time to Expiry
```{r}
n <- nrow(optionData) # Number of option positions

# Calculate time to expiry in terms of days

# If the Expiration.Time column Open and we need to minus 6.5/24 
# If the Expiration.Time column Close and we do not need to minus 6.5/24 

time_to_expiry_days <- rep(0, n)

for (i in 1:n) {
  t <- optionData$Expiration.Time[i]
  d <- optionData$Expiration.Date[i]
  
  if (t == "Open") {
    time_to_expiry_days[i] <- d - DATE - (6.5/24)
  } else {
    time_to_expiry_days[i] <- d - DATE
  }
}

# Add time_to_expiry to optionData
optionData$time_to_expiry_days <- time_to_expiry_days
optionData$time_to_expiry_years <- time_to_expiry_days / 365 # Annualized time_to_expiry
optionData$time_to_expiry_days
```

### Interest Rates 
Linear Interpolation of interest rate
Rn = R1 + (R2 - R1) / (t2 - t1) * (tn - t1)
```{r}
rates <- rep(0, n)

# Rates for each position
for (i in 1:n) {
  tn <- optionData$time_to_expiry_days[i]
  
  # Search through yield curve to find appropriate parameters
  for (j in 1:nrow(yieldCurve)) {
    if (yieldCurve$days[j] >= tn) {
      t1 <- yieldCurve$days[j - 1]
      t2 <- yieldCurve$days[j]
      r1 <- yieldCurve$rate[j - 1]
      r2 <- yieldCurve$rate[j]
      break
    }
  }
  
  # Linear interpolation
  rates[i] <- r1 + (r2 - r1) * (tn - t1) / (t2 - t1)
}

# Add rates to optionData
optionData$Rate <- rates / 100 # rates in %
optionData$Rate
```

### Implied Volatility
```{r}
vol <- rep(0, n)

for (i in 1:n) {
  vol[i] <- GBSVolatility(price = optionData[i, "Price"], TypeFlag = "c", S = SPOT, X = optionData[i, "Strike.Price"], Time = optionData[i, "time_to_expiry_years"], r = optionData[i, "Rate"], b = optionData[i, "Rate"] - q)
}

optionData$Implied.volatility <- vol
optionData$Implied.volatility
```

### Compute d1 and d2
{equation}
R vectorization convenient
```{r}
X <- optionData$Strike.Price
r <- optionData$Rate
sig <- optionData$Implied.volatility
t <- optionData$time_to_expiry_years

optionData$d1 <- ((ln(SPOT/X) + (r - q + 0.5 * sig ^ 2) * t)) / (sig * sqrt(t))
optionData$d2 <- optionData$d1 - sig * sqrt(t)
```

### Greeks
{equations}
```{r}
X <- optionData$Strike.Price
Multiplier <- optionData$Multiplier
t <- optionData$time_to_expiry_years
r <- optionData$Rate
sig <- optionData$Implied.volatility
d1 <- optionData$d1
d2 <- optionData$d2

optionData$Delta <- Multiplier * exp(-q * t) * pnorm(d1)
optionData$Gamma <- Multiplier * exp(-q * t) * dnorm(d1) / (SPOT * sig * sqrt(t))
optionData$Theta <- Multiplier * {-0.5 * SPOT * sig * exp(-q * t) * dnorm(d1) / sqrt(t) - r * X * exp(-r * t) * pnorm(d2) + q * SPOT * exp(-q * t) * pnorm(d1)}
optionData$Vega  <- Multiplier * SPOT * exp(-q * t) * sqrt(t) * dnorm(d1)
optionData[, c("Strike.Price", "Quantity", "Implied.volatility", "Delta", "Gamma", "Theta", "Vega", "time_to_expiry_years")]
```

### Portfolio Greeks
linearity of greeks
```{r}
portfolioGreeks <- data.frame(row.names = "Portfolio",
                              Delta = sum(optionData$Delta * optionData$Quantity),
                              Gamma = sum(optionData$Gamma * optionData$Quantity),
                              Theta = sum(optionData$Theta * optionData$Quantity),
                              Vega = sum(optionData$Vega * optionData$Quantity))
portfolioGreeks
```

### Portfolio value as SPX value changes
```{r}
# Parameters
X <- optionData$Strike.Price
Quantity <- optionData$Quantity
Multiplier <- optionData$Multiplier
sig <- optionData$Implied.volatility
t <- optionData$time_to_expiry_years
r <- optionData$Rate

# Calculate portfolio value for SPX values ranging from -20% to +20% of spot price
value <- c()
for (s in as.integer(0.8 * SPOT):as.integer(1.2 * SPOT)) {
  
  # Calculate value for each option position as spot price changes
  # Since parameters are all vectorized, value_part will also be a vector
  value_part <- Multiplier * Quantity * GBSOption(TypeFlag = "c", S = s, X = X, Time = t, r = r, b = r - q, sigma = sig)@price
  
  # Append portfolio value
  value <- c(value, sum(value_part))
}
```

### Data Visualization
```{r echo=FALSE}
priceChange <- data.frame(SPX = seq(-20, 20, 40 / (as.integer(1.2 * SPOT) - as.integer(0.8 * SPOT))), 
                          Value = value)

priceChange %>% 
  ggplot(aes(x=SPX, y=Value/1000000)) +
  geom_line() +
  ggtitle("Portfolio value as a function of % change in stock index") +
  xlab("% change in stock index") +
  ylab("Portfolio value (million $)") +
  scale_y_continuous(labels = comma)
```

## Indices Overview

### Load Data
```{r}
# Load data
SPX_history <- read.csv("data/SPX_History.csv")
VIX_history <- read.csv("data/VIX_History.csv")

SPX <- subset(SPX_history, select = c(Date,Close))
VIX <- subset(VIX_history, select = c(Date,Close))
SPX$Date <- as.Date(SPX$Date, format="%m/%d/%Y")
VIX$Date <- as.Date(VIX$Date, format="%m/%d/%Y")

# Merge dataframes
colnames(SPX)[colnames(SPX) == "Close"] <- "SPX"
colnames(VIX)[colnames(VIX) == "Close"] <- "VIX"
indexData <- merge(SPX, VIX, by = "Date")
indexData
```

### SPX Index
```{r echo=FALSE, out.width="100%"}
SPX %>%
  ggplot(aes(x=Date, y=SPX)) +
  geom_line() +
  ggtitle("Historical SPX Index Value") +
  xlab("Date") +
  ylab("SPX Value") +
  scale_y_continuous(labels = comma)
```

### VIX Index
```{r echo=FALSE, out.width="100%"}
VIX %>%
  ggplot(aes(x=Date, y=VIX)) +
  geom_line() +
  ggtitle("Historical VIX Index Value") +
  xlab("Date") +
  ylab("VIX Value") +
  scale_y_continuous(labels = comma)
```


## Modeling SPX and VIX Index

### Compute Index Log Return
```{r}
n_idx <- nrow(indexData)
indexData$SPX.Return[2:n_idx] <- log(indexData$SPX[2:n_idx] / indexData$SPX[1:(n_idx - 1)])
indexData$VIX.Return[2:n_idx] <- log(indexData$VIX[2:n_idx] / indexData$VIX[1:(n_idx - 1)])

# Get the most recent 1000 observations
idx_DATE <- which(indexData$Date == DATE)
returns <- indexData[(idx_DATE - 999):idx_DATE, c("Date", "SPX.Return", "VIX.Return")]
```

### Model Log Returns using GARCH
{equation}
```{r}
# Univariate GARCH specification prior to fitting
# Estimation of the parameters
spec1 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                    mean.model = list(armaOrder = c(0,0), include.mean = FALSE),
                    distribution.model = "norm")

spec2 <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
                    mean.model = list(armaOrder = c(1,0), include.mean = TRUE),
                    distribution.model = "norm")

# Fitting GARCH model
fitSPX <- ugarchfit(spec = spec1, data = returns[, "SPX.Return"])
fitVIX <- ugarchfit(spec = spec2, data = returns[, "VIX.Return"])

# GARCH coefficients
coef(fitSPX)
coef(fitVIX)
```

### Estimation of the DCC model
{equation}
```{r}
multiSpec <- multispec(c(spec1, spec2))
dccSpec <- dccspec(multiSpec, dccOrder = c(1,1), model = "DCC", distribution = "mvnorm")
fitDcc <- dccfit(dccSpec,returns[, c("SPX.Return", "VIX.Return")])

# DCC coefficients
coef(fitDcc)
```

### Store Coefficients
```{r}
SPXcoef <- fitSPX@fit$coef
SPXsigma <- fitSPX@fit$sigma
SPXz <- fitSPX@fit$z
VIXcoef <- fitVIX@fit$coef
VIXsigma <- fitVIX@fit$sigma
VIXz <- fitVIX@fit$z   
DCCcoef <- fitDcc@mfit$coef
DCCQ <- fitDcc@mfit$Q
```

### Conditional Mean, Variance, and Correlation for the next date
{equations}
```{r}
# Conditional mean
SPXmean <- 0
VIXmean <- VIXcoef["mu"] + VIXcoef["ar1"] * returns[1000, "VIX.Return"]


# Conditional variance
SPXvariance <- SPXcoef["omega"] + SPXcoef["alpha1"] * returns[1000, "SPX.Return"]^2 + SPXcoef["beta1"] * SPXsigma[1000]^2
VIXvariance <- VIXcoef["omega"] + VIXcoef["alpha1"] * returns[1000, "VIX.Return"]^2 + VIXcoef["beta1"] * VIXsigma[1000]^2

# Conditional correlation
rhobar <- (1/1000) * SPXz %*% VIXz
q11 <- 1.0 + DCCcoef["[Joint]dcca1"] * (SPXz[1000]^2 - 1) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][1,1] - 1)
q22 <- 1.0 + DCCcoef["[Joint]dcca1"] * (VIXz[1000]^2 - 1) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][2,2] - 1)
q12 <- rhobar + DCCcoef["[Joint]dcca1"] * (SPXz[1000] * VIXz[1000] - rhobar) + DCCcoef["[Joint]dccb1"] * (DCCQ[[1000]][1,2] - rhobar)
rho <- q12 / sqrt(q11 * q22)

# Cleaner output
names(VIXmean) <- NULL
names(SPXvariance) <- NULL
names(VIXvariance) <- NULL
rho <- rho[1, 1]

# Output
SPXmean
VIXmean
SPXvariance
VIXvariance
rho
```

### Portfolio Expected Change, Variance, and Delta-Normal VaR (Value at Risk)
{equation}
```{r}
delta <- portfolioGreeks$Delta
vega <- portfolioGreeks$Vega

# Portfolio expected change
portExpectedChange <- vega * (indexData[idx_DATE, "VIX"] / 100) * VIXmean

# Portfolio Variance
portVariance <- delta^2 * indexData[idx_DATE, "SPX"]^2 * SPXvariance + vega^2 * (indexData[idx_DATE, "VIX"] / 100)^2 * VIXvariance + delta * vega * indexData[idx_DATE, "SPX"] * (indexData[idx_DATE, "VIX"] / 100) * rho * sqrt(SPXvariance * VIXvariance)

# Portfolio standard deviation
portSD <- sqrt(portVariance)

# Portfolio delta-normal VaR
DNVaR <- -1.645 * (portExpectedChange - portSD)
names(DNVaR) <- NULL
DNVaR
```

## Filtered Historical Simulation
{equation}

## Filtered Historical Simulation VaR for 1-Day Holding Period

### Portfolio Value on Evaluation Date
```{r}
X <- optionData$Strike.Price
Quantity <- optionData$Quantity
Multiplier <- optionData$Multiplier
sig <- optionData$Implied.volatility
t <- optionData$time_to_expiry_years
r <- optionData$Rate

# Portfolio value on 10/30/2020
initPortValue <- sum(Multiplier * Quantity * GBSOption(TypeFlag = "c", S = SPOT, X = X, Time = t, r = r, b = r - q, sigma = sig)@price)
initPortValue
```

### Simulated Returns
Using parameter estimates from DCC model in previous section [DCC Estimates][Conditional Mean, Variance, and Correlation for the next date].
```{r}
SPX_simReturn <- SPXmean + sqrt(SPXvariance) * SPXz # Vector of 1000 simulated returns
VIX_simReturn <- VIXmean + sqrt(VIXvariance) * VIXz # Vector of 1000 simulated returns
```

### FHS Value at Risk
```{r}
newTau <- t - 3/365
newPortValue <- rep(0, 1000)
PL <- rep(0, 1000)

for (i in 1:1000) {
  
  # Calculate new SPX spot value using simulated return of SPX
  newSpot <- SPOT * exp(SPX_simReturn[i])
  
  # Calculate new volatility using simulated return of VIX
  newSigma <- sig * exp(VIX_simReturn[i])
  
  # Store new portfolio value
  newPortValue[i] <- sum(Multiplier * Quantity * GBSOption(TypeFlag = "c", S = newSpot, X = X, Time = newTau, r = r, b = r - q, sigma = newSigma)@price)
  
  # Store profit and loss
  PL[i] <- newPortValue[i] - initPortValue
}

FHSVaR <- -quantile(PL, 0.05)
FHSVaR
```


## Filtered Historical Simulation VaR for 1-Month Holding Period
Using parameter estimates from DCC model in previous section [DCC Estimates][Conditional Mean, Variance, and Correlation for the next date].
```{r}
X <- optionData$Strike.Price
Quantity <- optionData$Quantity
Multiplier <- optionData$Multiplier
sig <- optionData$Implied.volatility
t <- optionData$time_to_expiry_years
r <- optionData$Rate
```






